### 2024-11-23 paper/paper.qmd
Updated model set up of paper

---
title: "Number of COVID-19 Cases in Canada By Region, Age, and Sex"
subtitle: "Predicting the Factors that Influence the Number of COVID-19 Cases and Deaths"
author: 
  - Bruce Zhang
thanks: "Code and data are available at: [https://github.com/brucejczhang/covid_data](https://github.com/brucejczhang/covid_data)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)

# Load the dataset
analysis_data <- read.csv(here("data/02-analysis_data/analysis_data.csv"))
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR] to analyze the data and to create graphs and models. Th packages that were used include tidyverse [@tidyverse], tidyr [@tidyr], dplyr [@dplyr], and rstanarm [@rstanarm]. Our data [@covid_data] was obtained from COVerAGE-DB. The data anlysis was conducted based on the guidance of @tellingstories.

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

# Load necessary libraries
library(dplyr)
library(knitr)

# Calculate Case Fatality Rate (CFR) and summarize the data
summary_stats <- analysis_data %>%
  mutate(
    CFR = Deaths / (Cases + 1e-6) * 100 # Calculate Case Fatality Rate (percentage)
  ) %>%
  group_by(Region, Sex) %>% # Group by Region and Sex
  summarise(
    Mean_Age = mean(Age, na.rm = TRUE),             # Average age
    Total_Cases = sum(Cases, na.rm = TRUE),         # Total cases
    Total_Deaths = sum(Deaths, na.rm = TRUE),       # Total deaths
    Mean_CFR = mean(CFR, na.rm = TRUE)              # Average Case Fatality Rate
  ) %>%
  ungroup() # Remove grouping for display purposes

# Display the summary table
kable(summary_stats, digits = 2, col.names = c(
  "Region", "Sex", "Mean Age", "Total Cases", "Total Deaths", "Mean CFR (%)"
))

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

A logistic model has been selected as part of this analysis to examine the relationship between region, sex, age, and the number of cases and deaths of COVID-19.

The logistic regression model predicts the probability of a region or individual being classified as "high risk" based on key predictors. In this case, "high risk" is defined as regions where the death rate (CFR) exceeds the median death rate across the dataset.

The CFR is calculated as: 
\begin{align}
\text{CFR} = \frac{\text{Deaths}}{\text{Cases} + 10^{-6}} \times 100
\end{align}

The \textbf{box-and-whisker plot} visualizes the distribution of predicted probabilities (\( p \)) across different \textbf{age intervals}. Each box represents the interquartile range (IQR) of predicted probabilities, with the line inside the box indicating the median predicted probability for that age interval. Whiskers extend to 1.5 times the IQR, and any points outside this range are considered outliers. This visualization highlights the variability of predicted probabilities within each age group.

The logistic regression model equation is:
\begin{align}
\text{logit}(p) = \ln\left(\frac{p}{1 - p}\right) = \beta_0 + \beta_1 \cdot \text{Tests} + \beta_2 \cdot \text{Age Interval} + \epsilon
\end{align}

where:
\begin{itemize}
    \item $p$: Predicted probability of being high risk ($p = \text{Predicted Probability}$).
    \item $\text{logit}(p)$: Log-odds of the high-risk classification.
    \item $\beta_0$: Intercept of the model.
    \item $\beta_1, \beta_2$: Coefficients for the predictors (e.g., number of tests, age interval).
    \item $\epsilon$: Error term.
\end{itemize}

This model explains how testing rates and age intervals affect the likelihood of a region or demographic being classified as high risk. The box-and-whisker plot shows the variability of predicted probabilities across age groups, helping to identify specific age intervals with consistently higher or lower risks.

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
set.seed(21)

# Step 1: Filter for Urban and Rural Regions Only
filtered_data <- analysis_data %>%
  mutate(
    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
  )

# Step 2: Logistic Regression Model
# Fit a logistic regression model
logistic_model <- glm(
  High_Risk ~ Tests + Region + Age,
  family = binomial(),
  data = filtered_data
)

# Summary of the model
summary(logistic_model)

# Step 3: Add Predictions to the Dataset
filtered_data <- filtered_data %>%
  mutate(Predicted_Probability = predict(logistic_model, type = "response"))

# Step 4: Visualize the Results
# Plot predicted probabilities by region
ggplot(filtered_data, aes(x = Region, y = Predicted_Probability, fill = Region)) +
  geom_boxplot(alpha = 0.6) +
  labs(
    title = "Predicted Probability of High Risk by Region (Urban vs Rural)",
    x = "Region",
    y = "Predicted Probability of High Risk"
  ) +
  theme_minimal()

```

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
set.seed(21)

# Step 1: Filter for only 'f' and 'm', and define High Risk
filtered_data <- analysis_data %>%
  mutate(
    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
  )

# Step 2: Logistic Regression Model
# Fit a logistic regression model
logistic_model <- glm(
  High_Risk ~ Tests + Sex + Age,
  family = binomial(),
  data = filtered_data
)

# Summary of the model
summary(logistic_model)

# Step 3: Add Predictions to the Dataset
filtered_data <- filtered_data %>%
  mutate(Predicted_Probability = predict(logistic_model, type = "response"))

# Step 4: Visualize the Results
# Plot predicted probabilities by sex
ggplot(filtered_data, aes(x = Sex, y = Predicted_Probability, fill = Sex)) +
  geom_boxplot(alpha = 0.6) +
  labs(
    title = "Predicted Probability of High Risk by Sex",
    x = "Sex",
    y = "Predicted Probability of High Risk"
  ) +
  theme_minimal()

```

```{r}
# Load necessary libraries
library(dplyr)
library(ggplot2)
set.seed(21)

# Step 1: Create Age Intervals and Define High Risk
filtered_data <- analysis_data %>%
  mutate(
    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE), # Create age intervals
    Death_Rate = Deaths / (Cases + 1e-6), # Calculate death rate (avoid division by zero)
    High_Risk = ifelse(Death_Rate > median(Death_Rate, na.rm = TRUE), 1, 0) # Classify high-risk regions
  ) %>%
  filter(!is.na(Age_Interval)) # Remove any rows with missing age intervals

# Step 2: Logistic Regression Model
# Fit a logistic regression model
logistic_model <- glm(
  High_Risk ~ Tests + Age_Interval,
  family = binomial(),
  data = filtered_data
)

# Summary of the model
summary(logistic_model)

# Step 3: Add Predictions to the Dataset
filtered_data <- filtered_data %>%
  mutate(Predicted_Probability = predict(logistic_model, type = "response"))

# Step 4: Visualize the Results
# Plot predicted probabilities by age interval
ggplot(filtered_data, aes(x = Age_Interval, y = Predicted_Probability, fill = Age_Interval)) +
  geom_boxplot(alpha = 0.6) +
  labs(
    title = "Predicted Probability of High Risk by Age Interval",
    x = "Age Interval",
    y = "Predicted Probability of High Risk"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # Rotate x-axis labels for readability

```

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-country
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary library
library(ggplot2)

# Create a bar plot of total cases per country
ggplot(analysis_data, aes(x = Country, y = Cases, fill = Country)) +
  geom_bar(stat = "identity", color = "lightblue", alpha = 0.
)

```

```{r}
#| echo: false
#| eval: true
#| label: cases-over-time
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)
library(lubridate) # For date handling

# Filter data to include all regions
filtered_data <- analysis_data %>%
  filter(Region == "All") %>%
  mutate(
    # Ensure the Date column is in proper Date format
    Date = as.Date(Date, format = "%Y-%m-%d"),
    # Extract the year and month to group by
    Month = floor_date(Date, unit = "month")
  )

# Create the scatterplot with aggregated monthly cases
ggplot(filtered_data, aes(x = Date, y = Cases, color = Country, group = Country)) +
  geom_point(alpha = 0.7) +   # Add scatter points
  geom_line(size = 1) +       # Add lines connecting points for each country
  labs(
    title = "Monthly COVID Cases in Canada (Urban Region)",
    x = "Month",
    y = "Total Number of Cases",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )

```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-age
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries
selected_countries <- c("Canada")
filtered_data <- analysis_data %>%
  filter(Country %in% selected_countries) %>%
  mutate(
    # Create age intervals
    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE)
  )

# Aggregate the number of cases by country and age interval
age_case_summary <- filtered_data %>%
  group_by(Country, Age_Interval) %>%
  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>%
  ungroup()

# Create the faceted histogram
ggplot(age_case_summary, aes(x = Age_Interval, y = Total_Cases, fill = Country)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  facet_wrap(~Country, scales = "free_y") + # Create a grid of histograms for each country
  labs(
    title = "COVID Cases by Age Interval in Selected Countries",
    x = "Age Interval",
    y = "Number of Cases",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "grey80"),
    legend.position = "none"
  )

```

```{r}
#| echo: false
#| eval: true
#| label: deaths-by-age
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries
selected_countries <- c("Canada")
filtered_data <- analysis_data %>%
  filter(Country %in% selected_countries) %>%
  mutate(
    # Create age intervals
    Age_Interval = cut(Age, breaks = seq(0, 100, by = 10), right = FALSE, include.lowest = TRUE)
  )

# Aggregate the number of deaths by country and age interval
age_death_summary <- filtered_data %>%
  group_by(Country, Age_Interval) %>%
  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Use Deaths column
  ungroup()

# Create the faceted histogram for deaths
ggplot(age_death_summary, aes(x = Age_Interval, y = Total_Deaths, fill = Country)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") +
  facet_wrap(~Country, scales = "free_y") + # Create a grid of histograms for each country
  labs(
    title = "COVID Deaths by Age Interval in Selected Countries",
    x = "Age Interval",
    y = "Number of Deaths",
    fill = "Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major = element_line(color = "grey80"),
    legend.position = "none"
  )

```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-sex
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by gender and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
gender_case_summary <- analysis_data %>%
  filter(Country %in% selected_countries, Sex %in% c("f", "m")) %>% # Filter for selected countries and genders
  group_by(Country, Sex) %>%
  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>% # Summarize total cases by gender and country
  ungroup()

# Create the faceted bar graph
ggplot(gender_case_summary, aes(x = Sex, y = Total_Cases, fill = Sex)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  scale_fill_manual(values = c("f" = "pink", "m" = "blue")) + # Optional: Gender-specific colors
  labs(
    title = "COVID Cases by Gender in Selected Countries",
    x = "Gender",
    y = "Number of Cases",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```

```{r}
#| echo: false
#| eval: true
#| label: deaths-by-sex
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by gender and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
gender_death_summary <- analysis_data %>%
  filter(Country %in% selected_countries, Sex %in% c("f", "m")) %>% # Filter for selected countries and genders
  group_by(Country, Sex) %>%
  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Summarize total deaths by gender and country
  ungroup()

# Create the faceted bar graph
ggplot(gender_death_summary, aes(x = Sex, y = Total_Deaths, fill = Sex)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  scale_fill_manual(values = c("f" = "pink", "m" = "blue")) + # Optional: Gender-specific colors
  labs(
    title = "COVID Deaths by Gender in Selected Countries",
    x = "Gender",
    y = "Number of Deaths",
    fill = "Gender"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```

```{r}
#| echo: false
#| eval: true
#| label: cases-by-region
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by region and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
region_case_summary <- analysis_data %>%
  filter(Country %in% selected_countries) %>% # Filter for selected countries
  group_by(Country, Region) %>%
  summarise(Total_Cases = sum(Cases, na.rm = TRUE)) %>% # Summarize total cases by region and country
  ungroup()

# Create the faceted bar graph
ggplot(region_case_summary, aes(x = Region, y = Total_Cases, fill = Region)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  labs(
    title = "COVID Cases by Region in Selected Countries",
    x = "Region",
    y = "Number of Cases",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Rotate x-axis labels for readability
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```

```{r}
#| echo: false
#| eval: true
#| label: deaths-by-region
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Filter data for the selected 5 countries and group by region and country
selected_countries <- c("Afghanistan", "Brazil", "Canada", "Denmark", "Egypt")
region_death_summary <- analysis_data %>%
  filter(Country %in% selected_countries) %>% # Filter for selected countries
  group_by(Country, Region) %>%
  summarise(Total_Deaths = sum(Deaths, na.rm = TRUE)) %>% # Summarize total deaths by region and country
  ungroup()

# Create the faceted bar graph
ggplot(region_death_summary, aes(x = Region, y = Total_Deaths, fill = Region)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "black") + # Bar plot with totals
  facet_wrap(~Country, scales = "free_y") + # Facet by country
  labs(
    title = "COVID Deaths by Region in Selected Countries",
    x = "Region",
    y = "Number of Deaths",
    fill = "Region"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Rotate x-axis labels for readability
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    panel.spacing = unit(1, "lines") # Increase spacing between facets
  )

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

```{r}

```

## Diagnostics

```{r}

```

\newpage

# References